---
title: "identifying risky bank loans using C5.0 decision trees"
author: "Mastan Rao Parimi"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#exploring and preparing the data
data loading
```{r}
credit <- read.csv("credit.csv")
credit$default <- as.factor(credit$default)
```
summary
```{r}
str(credit)
```
table() output for a couple of loan features that seem likely to predict a default.
```{r}
table(credit$checking_balance)
```
```{r}
table(credit$savings_balance)
```
```{r}
summary(credit$months_loan_duration)
```
```{r}
summary(credit$amount)
```
The default vector indicates whether the loan applicant was unable to meet the
agreed payment terms and went into default. A total of 30 percent of the loans in
this dataset went into default:
```{r}
table(credit$default)
```
#Data preparation – creating random training and test datasets
```{r}
set.seed(123)
train_sample <- sample(1000, 900)
```
By using this vector to select rows from the credit data, we can split it into 
the 90 percent training and 10 percent test datasets we desired
```{r}
credit_train <- credit[train_sample, ]
credit_test <- credit[-train_sample, ]
```
should have about 30 percent of defaulted loans in each of the datasets:
```{r}
prop.table(table(credit_train$default))
prop.table(table(credit_test$default))
```
#training a model on the data
```{r}
library(C50)
```
Training the Model using c.50 function
```{r}
credit_model <- C5.0(credit_train[-17], credit_train$default)
```
call model
```{r}
credit_model
```
To see the tree's decisions, we can call the summary() function on the model:
```{r}
summary(credit_model)
```
#evaluating model performance
predicting using test data
```{r}
credit_pred <- predict(credit_model, credit_test)
```
```{r}
library(gmodels)
CrossTable(credit_test$default, credit_pred,
prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE,
dnn = c('actual default', 'predicted default'))
```
#improving model performance
```{r}
credit_boost10 <- C5.0(credit_train[-17], credit_train$default,
trials = 10)
```
resulting model
```{r}
credit_boost10
```
summary of the model
```{r}
summary(credit_boost10)
```
predicting using test data
```{r}
credit_boost_pred10 <- predict(credit_boost10, credit_test)
CrossTable(credit_test$default, credit_boost_pred10,
prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE,
dnn = c('actual default', 'predicted default'))
```
cost matrix
```{r}
matrix_dimension <- list(c("1", "2"), c("1", "2"))
names(matrix_dimension) <- c("predicted", "actual")
matrix_dimension
```
Suppose we believe that a loan default costs the bank four times as much as a 
missed opportunity. Our penalty values could then be defined as:
```{r}
error_cost <- matrix(c(0, 1, 4, 0), nrow = 2,
dimnames = matrix_dimension)
error_cost
```
As defined by this matrix, there is no cost assigned when the algorithm
classifies a no or yes correctly, but a false negative has a cost of 4 versus a 
false positive's cost of 1. To see how this impacts classification, let's apply 
it to our decision tree using the costs parameter of the C5.0() function. We'll 
otherwise use the same steps as we did earlier:
```{r} 
credit_cost <- C5.0(credit_train[-17], credit_train$default,
costs = error_cost)
credit_cost_pred <- predict(credit_cost, credit_test)
CrossTable(credit_test$default, credit_cost_pred,
prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE,
dnn = c('actual default', 'predicted default'))
```